A Theory of Bidirectional Type Synthesis for Simple Types

### 2023-06-29 tex/sec2-outline.tex
[ tex ] Done with S3 for now

\begin{figure}
  \centering
  \small
  \judgbox{|- \isTerm{t}^\dir{d}}{A raw term~$\isTerm{t}$ is in mode~$d$}
  \begin{mathpar}
    \inferrule{x \in V}{V |- \isTerm{x}^{\syn}}\,\SynRule{Var}
    \and
    \inferrule{V |- \isTerm{t}^{\chk}}{V |- (\isTerm{t \annotate A})^{\syn}}\,\SynRule{Anno}
    \and
    \inferrule{V |- \isTerm{t}^{\syn}}{V |- \isTerm{t}^{\chk}}\,\ChkRule{Sub}
    \\
    \inferrule{V, x |- \isTerm{t}^{\chk}}
    {V |- (\isTerm{\lam{x}t})^{\chk}}\,\ChkRule{Abs}
    \and
    \inferrule{V |- \isTerm{t}^{\syn} \\ V |- \isTerm{u}^{\chk}}
    {V |- (\isTerm{t\;u})^{\syn}}\,\SynRule{App}
  \end{mathpar}
  \caption{Mode derivations for simply typed $\lambda$-calculus}
  \label{fig:STLC-mode-derivations}
\end{figure}

@@

\begin{figure}
  \begin{tikzpicture}
  \matrix (m) [matrix of math nodes, row sep=3em, column sep=10em]
  {
    \boxed{\erase{\Gamma} \vdash \isTerm{t}}   & \boxed{\erase{\Gamma} \vdash \isTerm{t}^{\syn}} \\
    \boxed{\Gamma \vdash \isTerm{t} : A} & \boxed{\Gamma \vdash \isTerm{t} \syn A} \\
  };
  \path[->, line width=0.7pt]
      (m-1-1) edge node [above] {mode preprocessing} (m-1-2)
      (m-1-2) edge node [right,align=center] {bidirectional\\type synthesis} (m-2-2)
      (m-2-1) edge[bend right=8] node [below] {completeness} (m-2-2)
      (m-2-2) edge[bend right=8] node [above] {soundness}    (m-2-1);
  \path[dashed,->,line width=0.7pt]
      (m-1-1) edge node [left,align=center]  {type\\synthesis} (m-2-1);
  \end{tikzpicture}
  \caption{An implementation of type synthesis via bidirectional typing}
\end{figure}

>>>>

\begin{figure}
  \centering
  \small
  \judgbox{V |- \isTerm{t}^\dir{d}}{A raw term~$\isTerm{t}$ is in mode~$d$ with free variables in $V$}
  \begin{mathpar}
    \inferrule{x \in V}{V |- \isTerm{x}^{\syn}}\,\SynRule{Var}
    \and
    \inferrule{V |- \isTerm{t}^{\chk}}{V |- (\isTerm{t \annotate A})^{\syn}}\,\SynRule{Anno}
    \and
    \inferrule{V |- \isTerm{t}^{\syn}}{V |- \isTerm{t}^{\chk}}\,\ChkRule{Sub}
    \\
    \inferrule{V, x |- \isTerm{t}^{\chk}}
    {V |- (\isTerm{\lam{x}t})^{\chk}}\,\ChkRule{Abs}
    \and
    \inferrule{V |- \isTerm{t}^{\syn} \\ V |- \isTerm{u}^{\chk}}
    {V |- (\isTerm{t\;u})^{\syn}}\,\SynRule{App}
  \end{mathpar}
  \caption{Mode derivations for simply typed $\lambda$-calculus}
  \label{fig:STLC-mode-derivations}
\end{figure}

@@

\begin{figure}
  \small
  \begin{tikzpicture}
  \matrix (m) [matrix of math nodes, row sep=3em, column sep=10em]
  {
    \boxed{\erase{\Gamma} \vdash \isTerm{t}}   & \boxed{\erase{\Gamma} \vdash \isTerm{t}^{\syn}} \\
    \boxed{\Gamma \vdash \isTerm{t} : A} & \boxed{\Gamma \vdash \isTerm{t} \syn A} \\
  };
  \path[->, line width=0.5pt]
      (m-1-1) edge node [above] {mode preprocessing} (m-1-2)
      (m-1-2) edge node [right,align=center] {bidirectional\\type synthesis} (m-2-2)
      (m-2-1) edge[bend right=8] node [below] {completeness} (m-2-2)
      (m-2-2) edge[bend right=8] node [above] {soundness}    (m-2-1);
  \path[dashed,->,line width=0.5pt]
      (m-1-1) edge node [left,align=center]  {type\\synthesis} (m-2-1);
  \end{tikzpicture}
  \caption{An implementation of type synthesis via bidirectional typing}
\end{figure}
